---
title: "Prediction using Supervised ML"
author: "Priyanka Puppala"
date: "Last Updated: `r format(Sys.time(), '%d, %B, %Y at %H:%M')`"
output: rmdformats::readthedown
--- 
### Graduate Rotational Internship Program The Sparks Foundation
### Data Science & Business Analytics Tasks
### Prediction using Supervised ML

# Introduction
* Linear Regression with R Programming (R-studio)

* In this project the we can see how a Simple Linear Regression can be used for 
prediction for the percentage of an student based on the no. of study hours.

* It is implementing using R programming -rmarkdown report.

* We will start with simple linear regression involving two variables.

# Problem Statement

* In this we need to predict the percentage of marks based on the number of 
hours they study.
* Here we are using  two variables Hours and  Scores for the regression. 

# Data Collection

* Collecting the student data from url
* Reading the data into csv file 

```{r,,warning=FALSE, message=FALSE }
library(readr)
library(tidyverse)
library(tidymodels)
library(ggplot2)
student_scoresURL <- "https://raw.githubusercontent.com/AdiPersonalWorks/Random/master/student_scores%20-%20student_scores.csv"

student_scores <- read.csv(file = student_scoresURL, 
                  header = TRUE,
                     sep = ",",
                     stringsAsFactors = FALSE)
student_scores
```
* writing/saving the student_scores file into local directory
```{r , warning=FALSE, message=FALSE}

write.table(student_scores, file = "student_scores.csv ",
            row.names = FALSE,
            col.names = TRUE,
            sep = "\t",
            quote = FALSE)



```

* Checking Basic Information about the student_scores data
* Checking summary and structure of the data 
```{r , warning=FALSE, message=FALSE}
head(student_scores)
dim(student_scores)

str(student_scores) 

summary(student_scores) 
```
# Data Preparation
```{r , warning=FALSE, message=FALSE}
library(caTools)
set <- sample(2, nrow(student_scores), 
              replace = TRUE, 
              prob = c(0.7, 0.3))
train <- student_scores[set==1,]
train

test <- student_scores[set==2,]
test
```
## Data Cleaning

* Checking for Missing Values/NA/NAN
```{r , warning=FALSE, message=FALSE}
library(DataExplorer)
sum(is.na(train))
sum(is.na(test))
plot_missing(train)
```
# Exploratory Data Analysis (EDA)
* describe - can computes the statistics of all numerical variables 
```{r, warning=FALSE, message=FALSE}
library(Hmisc)
describe(train)
describe(test)

```
* Two continuous variables
```{r, fig.width=8, fig.height=5}
library(ggplot2)

q <- ggplot(data = train, aes(x = Hours, y = Scores))+
  geom_line(colour = "skyblue") +
  geom_point(colour = "blue")
q

ggplot(data=train,aes(x=Hours,y=Scores)) + 
  geom_bar(stat ='identity',aes(fill=Scores))+
  coord_flip() + 
  theme_grey() + 
  scale_fill_gradient(name="Score Level")+
  labs(title = 'Scores according to hours',
       y='Score per hour',x='Hours of study')+ 
  geom_hline(yintercept = mean(student_scores$Scores),size = 1, color = 'red')


```
```{r, fig.width=8, fig.height=5, message=FALSE, warning=FALSE}
library(plotly)
plot_ly(train, x = ~Hours, y = ~Scores, type = 'bar', mode = 'markers',marker = list(color = "DARKGREEN", opacity = 0.5), size = 4) %>%  
  layout(title = 'Scores according to hours', 
                       yaxis = list(title = 'Scores'), 
                       xaxis = list(title = 'Hours of study') )



```


# Modelling
* Simple Linear Regression
```{r , warning=FALSE, message=FALSE}


model <- lm( Scores~  Hours,  data=train) #model building
model
```
* Summary of model
```{r , warning=FALSE, message=FALSE}
summary(model) 

#plot
plot(Scores~  Hours,  train, pch =4, frame = FALSE, col  = "red")
abline(model, col = "blue")
```


# Model Prediction
```{r , warning=FALSE, message=FALSE}
# Prediction
cof <- coef(summary(model))
cof

pred  <- predict(model,train)
pred

train_pred <- data.frame(train,pred)
train_pred

```


## Comparing actual values with Predicted values
```{r , warning=FALSE, message=FALSE}
Scores_comp <- data.frame(Actual  = train$Scores, Predcited = train_pred$pred)
Scores_comp
```


### Actual vs Predicted Graph
```{r , warning=FALSE, message=FALSE, fig.width=8, fig.height=5}
Actual <- c(Scores_comp$Actual)
Predcited <- c(Scores_comp$Predcited)
plot(Actual, type = "b", frame = FALSE, pch = 10, 
     col = "#4AC6B7", xlab = "x", ylab = "y")

lines(Predcited, pch = 10, col = "#C61951", type = "b", lty = 2)

legend("topleft", legend=c("Actual", "Predicted"), col=c("#4AC6B7", "#C61951"), 
       pch = 10, lty = 1:1, cex=0.6)
```

# Model Diagnostics
```{r , warning=FALSE, message=FALSE, fig.width=8, fig.height=5}

par(mfrow=c(2,2))

plot(model)


```
## Test data Evoluation
```{r , warning=FALSE, message=FALSE}

pred  <- predict(model,test)
tibble(pred)

test_pred <- data.frame(test,pred)
test_pred

```
### Actual vs Predicted graph

```{r , warning=FALSE, message=FALSE}
Scores_comp2 <- data.frame(Actual  = test$Scores, Predcited = test_pred$pred)
Scores_comp2
```
# The predicted score if a student studies for 9.25 hrs/ day  
* Predicting score if a student studies for 9.25 hrs/day
```{r , warning=FALSE, message=FALSE}
pred_hour  <-  predict(model, data.frame(Hours=9.25)) 
pred_hour

#(or)
coef(model)[1] + 9.25*coef(model)[2] #Mathematical form
```

```{r , include = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
train_pred_new <-  train_pred %>% add_row(Hours=9.25, Scores= pred_hour,
                                          pred =pred_hour)
train_pred_new
```


```{r , warning=FALSE, message=FALSE, fig.width=8, fig.height=5}
library(highcharter)

hc <- train_pred_new %>% group_by(Scores, Hours) %>%
  hchart('column', hcaes(x = 'Hours', y = 'Scores') ) 

hc
```